---
title: "CONN_LMM_PLOT"
output: html_document
date: "2025-07-01"
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)

# Load Libraries
library(dplyr) 
library(ggplot2)
# library(devtools)
# library(ggpmisc)
library(lme4)
#library(optimx)
library(nlme)
#library(growthcurver)
library(tidyr)
library(viridis)
library(RColorBrewer)
library(forcats)

```

```{r Organize the data into long format}
# set the directory
setwd("/Users/quanz/Documents/UM/Projects/Connectivity/LMM")


# file <- "CONN_LMM_2025.csv"
# file <- "CONN_Mastersheet_OAs.csv"
# June 25 updated
file <- "CONN_Mastersheet_2025_7.9.csv"
CONN_LMM <- read.csv(file)


CONN_LMM_long <- CONN_LMM %>%
  pivot_longer(
    cols = -c(Type, ID, SubNum, Age, between_age, Btwn_cAge, With_cAge), 
    names_to = c("ROI", "measure"), 
    names_sep = "_"
  ) %>%
  pivot_wider(
    names_from = measure,
    values_from = value
  ) %>%
  arrange(ROI)

CONN_LMM_long_old_combined <- CONN_LMM_long %>%
  filter(Type %in% c("W1", "W2"))
```


## Including Plots

You can also embed plots, for example:

```{r making the CONN_LMM_long_combined dataset for plot}
CONN_LMM_long_old <- CONN_LMM_long %>%
  filter(Type != "Young")

within_avg <- CONN_LMM_long_old %>%
  group_by(SubNum, Age) %>%  # Group by Subject and Age
  summarize(within = mean(within, na.rm = TRUE), .groups = "drop") %>%
  mutate(ROI = "Overall Mean")  # Assign a label for overall mean

within_avg <- within_avg %>%
  left_join(
    CONN_LMM_long_old %>% 
      filter(ROI == "DMN") %>% 
      select(SubNum, Type, ID, between_age, Btwn_cAge, With_cAge),
    by = "SubNum"
  ) %>%
  mutate(ROI = "Overall Mean")  # Keep "Overall Mean" as a label

CONN_LMM_long_old_combined <- bind_rows(CONN_LMM_long_old, within_avg)

roi_order <- c("DMN", "SMN", "SAL", "DAN", "FP", "LAN", "VIS", "CRB", "Overall Mean")

CONN_LMM_long_old_combined <- CONN_LMM_long_old_combined %>%
  filter(ROI %in% roi_order)

CONN_LMM_long_old_combined$ROI <- factor(CONN_LMM_long_old_combined$ROI, levels = roi_order)
```


```{r Model 2 - within-network connectivity longitudinal}
# ```{r fit the lme for each network for the plots}
# List of unique ROIs (networks)
roi_list <- c("DMN", "SMN", "SAL", "DAN", "FP", "LAN", "VIS", "CRB")

# Initialize an empty list to store model results
network_fits <- list()

# Loop through each network and fit a separate LME model
for (roi in roi_list) {
  
  # Subset data for the current ROI
  subset_data <- CONN_LMM_long_old_combined %>% filter(ROI == roi)
  
  # Fit LME model for the current network
  model_w_network <- lme(within ~ Btwn_cAge + With_cAge, 
               random = ~ 1  | SubNum, 
               method = "ML", 
               data = subset_data)
  
  # Extract fixed effects (intercept and slope for With_cAge)
  intercept <- fixef(model_w_network)["(Intercept)"]
  slope <- fixef(model_w_network)["With_cAge"]
  
  # Store results in a list
  network_fits[[roi]] <- data.frame(ROI = roi, Intercept = intercept, Slope = slope)
}

# Convert the list to a single dataframe
network_fits_df <- bind_rows(network_fits)

# View the results
# print(network_fits_df)

# Create Age Sequence
age_seq <- seq(min(CONN_LMM_long_old_combined$Age), max(CONN_LMM_long_old_combined$Age), length.out = 100)

# Generate Predictions for Each Network
network_predictions <- network_fits_df %>%
  rowwise() %>%
  mutate(
    predictions = list(
      tibble(
        Age = age_seq,
        Within_Network_Connectivity = Intercept + Slope * (age_seq - 71.54), #age centered as the average age of older adults who completed both waves
        ROI_Network = ROI  # Rename to avoid conflict
      )
    )
  ) %>%
  unnest(predictions)  # Expand the nested dataframe

```

```{r Model 2 fit within-network connectivity for all networks plot}
# Generate predicted values for each network (from network_predictions_df)
roi_order <- c("DMN", "SMN", "SAL", "DAN", "FP", "LAN", "VIS", "CRB", "Overall Mean")
roi_colors <- c(brewer.pal(8, "Set2"), "#1f3c88")  # Set2 has 8 colors, adding black

network_predictions$ROI <- factor(network_predictions$ROI, levels = roi_order)

overall_p_value <- 0.0041  # The p-value from your model for Within_cAge

# Create a data frame with just the "Overall Mean" p-value
p_value_overall <- data.frame(
  ROI = "Overall Mean",  
  #x = max(CONN_LMM_long_old_combined$Age, na.rm = TRUE) - 2,  # Shift near right edge
  #y = min(CONN_LMM_long_old_combined$within, na.rm = TRUE) - 0.01,  # Shift to bottom
  p_label = sprintf("p = %.4f", overall_p_value)
)

age_seq <- seq(min(CONN_LMM_long_old_combined$Age), max(CONN_LMM_long_old_combined$Age), length.out = 100)

slope_overall <-  fixef(model_w)["With_cAge"]
intercept_overall <- fixef(model_w)["(Intercept)"]

mle_predicted <- data.frame(
  Age = age_seq,  # Use actual Age, not centered
  within = intercept_overall + slope_overall * (age_seq - 71.54),  # Adjust for centering
  ROI = "Overall Mean"
)

# Ensure the factor levels are applied correctly in all relevant dataframes
CONN_LMM_long_old_combined$ROI <- factor(CONN_LMM_long_old_combined$ROI, levels = roi_order)
network_predictions$ROI <- factor(network_predictions$ROI, levels = roi_order)
mle_predicted$ROI <- factor(mle_predicted$ROI, levels = roi_order)
p_value_overall$ROI <- factor(p_value_overall$ROI, levels = roi_order)

# Plot using Set2 color palette
p <- ggplot(CONN_LMM_long_old_combined, aes(x = Age, y = within, group = SubNum, color = ROI)) +
  geom_point(size = 2) +  # Slightly larger points for visibility
  geom_line(alpha = 0.6) +

  # Add regression lines for each network, now colored by ROI
  geom_line(data = network_predictions, 
            aes(x = Age, y = Within_Network_Connectivity, color = ROI, group = ROI), 
            inherit.aes = FALSE, size = 1.2) +  # Match colors to networks
  
  # Add the overall MLE regression line in red
  geom_line(data = mle_predicted, aes(x = Age, y = within), 
            inherit.aes = FALSE, color = "red", size = 1.2) +
  
  facet_wrap(~ROI, scales = "free_y", ncol = 3, drop = FALSE) +  
  scale_color_manual(values = roi_colors) +  # Use Set2 color palette
  theme_minimal(base_size = 12) +  # Increase base font size
  labs(x = "Age", y = "Within-network Connectivity") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", family = "Times New Roman"),  # Center title
    plot.title.position = "plot",
    axis.title.x = element_text(size = 14, face = "bold", family = "Times New Roman"),  # Bigger x-axis title
    axis.title.y = element_text(size = 14, face = "bold", family = "Times New Roman"),  # Bigger y-axis title
    axis.text = element_text(size = 10),  # Bigger tick labels
    strip.text = element_text(size = 12, face = "bold", family = "Times New Roman"),  
    legend.position = "none"  
  ) + 
  # Add p-value annotation for Overall Mean
  geom_text(data = p_value_overall, 
          aes(x = max(CONN_LMM_long_old_combined$Age) - 5,  # Move it further right
              y = min(CONN_LMM_long_old_combined$within, na.rm = TRUE) + 0.55,  # Move it below the lowest point
              label = p_label), 
          inherit.aes = FALSE, size = 4, family = "Times New Roman", fontface = "italic", color = "red")

# --- Compute SAL p-value from the longitudinal model and annotate it ---
sal_dat <- CONN_LMM_long_old_combined %>% dplyr::filter(ROI == "SAL")

# Mirror your per-network longitudinal spec
m_sal <- nlme::lme(within ~ Btwn_cAge + With_cAge,
                   random = ~ 1 | SubNum,   # use SubNum if that's your column name
                   method = "ML",
                   data = sal_dat)

p_sal   <- summary(m_sal)$tTable["With_cAge", "p-value"]
lab_sal <- ifelse(p_sal < 1e-4, "p < .0001", sprintf("p = %.4f", p_sal))
col_sal <- ifelse(p_sal < 0.05, "red", "black")

# Anchor label inside the SAL facet (bottom-right) and add a little padding
annot_sal <- data.frame(
  ROI   = factor("SAL", levels = roi_order),
  Age   = Inf,
  within = -Inf,
  label = lab_sal
)

p <- p +
  geom_text(
    data = annot_sal,
    aes(x = Age, y = within, label = label),
    inherit.aes = FALSE,
    hjust = 1.02, vjust = -0.6,
    size = 4, family = "Times New Roman",
    fontface = "italic",
    color = col_sal
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.06)))


fp_dat <- CONN_LMM_long_old_combined %>% dplyr::filter(ROI == "FP")

m_fp <- nlme::lme(within ~ Btwn_cAge + With_cAge,
                  random = ~ 1 | SubNum,
                  method = "ML",
                  data = fp_dat)

p_fp   <- summary(m_fp)$tTable["With_cAge", "p-value"]
lab_fp <- ifelse(p_fp < 1e-4, "p < .0001", sprintf("p = %.4f", p_fp))
col_fp <- ifelse(p_fp < 0.05, "red", "black")

annot_fp <- data.frame(
  ROI    = factor("FP", levels = roi_order),
  Age    = Inf,
  within = -Inf,
  label  = lab_fp
)

p <- p +
  geom_text(
    data = annot_fp,
    aes(x = Age, y = within, label = label),
    inherit.aes = FALSE,
    hjust = 1.02, vjust = -0.6,
    size = 4, family = "Times New Roman",
    fontface = "italic",
    color = col_fp
  )


ggsave("Figure_5_long_within_network_change.tiff", plot = p, width = 17.6, height = 13.5, units = "cm", dpi = 300)

```

```{r Model 2 - Between-network}
# ```{r fit the lme for each network regarding between-network connectivity the plots}
# List of unique ROIs (networks)
roi_list <- c("DMN", "SMN", "SAL", "DAN", "FP", "LAN", "VIS", "CRB")

# Initialize an empty list to store model results
network_fits <- list()

# Loop through each network and fit a separate LME model
for (roi in roi_list) {
  
  # Subset data for the current ROI
  subset_data <- CONN_LMM_long_old_combined %>% filter(ROI == roi)
  
  # Fit LME model for the current network
  model_b_network <- lme(between ~ Btwn_cAge + With_cAge, 
               random = ~ 1  | SubNum, 
               method = "ML", 
               data = subset_data)
  
  # Extract fixed effects (intercept and slope for With_cAge)
  intercept_b_network <- fixef(model_b_network)["(Intercept)"]
  slope_b_network <- fixef(model_b_network)["With_cAge"]
  
  # Store results in a list
  network_fits[[roi]] <- data.frame(ROI = roi, Intercept_b_network = intercept_b_network, Slope_b_network = slope_b_network)
}

# Convert the list to a single dataframe
network_fits_df <- bind_rows(network_fits)

# View the results
# print(network_fits_df)

# Create Age Sequence
age_seq <- seq(min(CONN_LMM_long_old_combined$Age), max(CONN_LMM_long_old_combined$Age), length.out = 100)

# Generate Predictions for Each Network
network_predictions_b_network <- network_fits_df %>%
  rowwise() %>%
  mutate(
    predictions = list(
      tibble(
        Age = age_seq,
        Between_Network_Connectivity = Intercept_b_network + Slope_b_network * (age_seq - 71.54),
        ROI_Network = ROI  # Rename to avoid conflict
      )
    )
  ) %>%
  unnest(predictions)  # Expand the nested dataframe

```


```{r plot the Model 2 Between network connectivity for all the networks}
# Step 1: Calculate the average between-network connectivity for each subject & age
between_avg <- CONN_LMM_long_old %>%
  group_by(SubNum, Age) %>%  # Group by Subject and Age
  summarize(between = mean(between, na.rm = TRUE), .groups = "drop") %>%
  mutate(ROI = "Overall Mean")  # Assign a label for overall mean

# Step 2: Merge `between_avg` into the `Overall Mean` rows
CONN_LMM_long_old_combined <- CONN_LMM_long_old_combined %>%
  mutate(
    between = ifelse(ROI == "Overall Mean", 
                     between_avg$between[match(paste(SubNum, Age), 
                                               paste(between_avg$SubNum, between_avg$Age))], 
                     between)  # Keep original values for other ROIs
  )


```

```{r plot the between network connectivity for all the networks}
# Ensure ROI levels are correctly ordered
network_predictions_b_network$ROI <- factor(network_predictions_b_network$ROI, levels = roi_order)

# Set color scheme
roi_order <- c("DMN", "SMN", "SAL", "DAN", "FP", "LAN", "VIS", "CRB", "Overall Mean")
roi_colors <- c(brewer.pal(8, "Set2"), "#1f3c88")  # Set2 has 8 colors, adding black

# Create p-value annotation for Overall Mean
p_value_b_overall <- 0.1251
p_value_b_overall <- data.frame(
  ROI = "Overall Mean",  
  p_label_b = sprintf("p = %.4f", p_value_b_overall)
)

mle_predicted_between <- data.frame(
  Age = age_seq,  
  between = fixef(model_b)["(Intercept)"] + fixef(model_b)["With_cAge"] * (age_seq - 71.54),  
  ROI = "Overall Mean"
)

# Ensure the factor levels are applied correctly in all relevant dataframes
CONN_LMM_long_old_combined$ROI <- factor(CONN_LMM_long_old_combined$ROI, levels = roi_order)
network_predictions_b_network$ROI <- factor(network_predictions_b_network$ROI, levels = roi_order)
mle_predicted_between$ROI <- factor(mle_predicted_between$ROI, levels = roi_order)
p_value_b_overall$ROI <- factor(p_value_b_overall$ROI, levels = roi_order)

# **📊 Plot for Between-Network Connectivity**
between_network_plot <- ggplot(CONN_LMM_long_old_combined, aes(x = Age, y = between, group = SubNum, color = ROI)) +
  geom_point(size = 2) +  # Slightly larger points for visibility
  geom_line(alpha = 0.6) +  

  # Regression lines for each network (colored)
  geom_line(data = network_predictions_b_network, 
            aes(x = Age, y = Between_Network_Connectivity, color = ROI, group = ROI), 
            inherit.aes = FALSE, size = 1.2) +  

  # Overall Mean regression line in **red**
  geom_line(data = mle_predicted_between, aes(x = Age, y = between), 
            inherit.aes = FALSE, color = "black", size = 1.2) +

  # Facet by network
  facet_wrap(~ROI, scales = "free_y", ncol = 3, drop = FALSE) +  
  scale_color_manual(values = roi_colors) +  

  # Improve theme
  theme_minimal(base_size = 14) +  
  labs(x = "Age", y = "Between-network Connectivity") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", family = "Times New Roman"),  
    axis.title.x = element_text(size = 14, face = "bold", family = "Times New Roman"),  
    axis.title.y = element_text(size = 14, face = "bold", family = "Times New Roman"),  
    axis.text = element_text(size = 10),  
    strip.text = element_text(size = 12, face = "bold", family = "Times New Roman"),  
    legend.position = "none"
  ) +  

  # P-value annotation for Overall Mean
  geom_text(data = p_value_b_overall, 
          aes(x = max(CONN_LMM_long_old_combined$Age) - 5,  
              y = min(CONN_LMM_long_old_combined$between, na.rm = TRUE) + 0.4,  
              label = p_label_b), 
          inherit.aes = FALSE, size = 4, fontface = "italic", color = "black")

ggsave("Fig.6_long_between_network_change.tiff", plot = between_network_plot, width = 17.6, height = 13.5, units = "cm", dpi = 300)
```

```{r Model 2 - Segregation}
# List of unique ROIs (networks)
roi_list <- c("DMN", "SMN", "SAL", "DAN", "FP", "LAN", "VIS", "CRB")

# Initialize an empty list to store model results
network_fits <- list()

# Loop through each network and fit a separate LME model
for (roi in roi_list) {
  
  # Subset data for the current ROI
  subset_data <- CONN_LMM_long_old_combined %>% filter(ROI == roi)
  
  # Fit LME model for the current network
  model_s_network <- lme(seg ~ Btwn_cAge + With_cAge, 
               random = ~ 1  | SubNum, 
               method = "ML", 
               data = subset_data)
  
  # Extract fixed effects (intercept and slope for With_cAge)
  intercept_s_network <- fixef(model_s_network)["(Intercept)"]
  slope_s_network <- fixef(model_s_network)["With_cAge"]
  
  # Store results in a list
  network_fits[[roi]] <- data.frame(ROI = roi, Intercept_s_network = intercept_s_network, Slope_s_network = slope_s_network)
}

# Convert the list to a single dataframe
network_fits_df <- bind_rows(network_fits)

# View the results
# print(network_fits_df)

# Create Age Sequence
age_seq <- seq(min(CONN_LMM_long_old_combined$Age), max(CONN_LMM_long_old_combined$Age), length.out = 100)

# Generate Predictions for Each Network
network_predictions_s_network <- network_fits_df %>%
  rowwise() %>%
  mutate(
    predictions = list(
      tibble(
        Age = age_seq,
        Segregation = Intercept_s_network + Slope_s_network * (age_seq - 71.54),
        ROI_Network = ROI  # Rename to avoid conflict
      )
    )
  ) %>%
  unnest(predictions)  # Expand the nested dataframe


```

```{r plot Model 2 seg for all the networks}
# Step 1: Calculate the average between-network connectivity for each subject & age
seg_avg <- CONN_LMM_long_old %>%
  group_by(SubNum, Age) %>%  # Group by Subject and Age
  summarize(seg = mean(seg, na.rm = TRUE), .groups = "drop") %>%
  mutate(ROI = "Overall Mean")  # Assign a label for overall mean

# Step 2: Merge `between_avg` into the `Overall Mean` rows
CONN_LMM_long_old_combined <- CONN_LMM_long_old_combined %>%
  mutate(
    seg = ifelse(ROI == "Overall Mean", 
                     seg_avg$seg[match(paste(SubNum, Age), 
                                               paste(seg_avg$SubNum, seg_avg$Age))], 
                     seg)  # Keep original values for other ROIs
  )
```

```{r plot the between network connectivity for all the networks}
# Ensure ROI levels are correctly ordered
network_predictions_s_network$ROI <- factor(network_predictions_s_network$ROI, levels = roi_order)

# Set color scheme
roi_order <- c("DMN", "SMN", "SAL", "DAN", "FP", "LAN", "VIS", "CRB", "Overall Mean")
roi_colors <- c(brewer.pal(8, "Set2"), "#1f3c88")  # Set2 has 8 colors, adding black

# Create p-value annotation for Overall Mean
p_value_s_overall <- 0.7490
p_value_s_overall <- data.frame(
  ROI = "Overall Mean",  
  p_label_s = sprintf("p = %.4f", p_value_s_overall)
)

mle_predicted_seg <- data.frame(
  Age = age_seq,  
  seg = fixef(model_s)["(Intercept)"] + fixef(model_s)["With_cAge"] * (age_seq - 71.54),  
  ROI = "Overall Mean"
)

# Ensure the factor levels are applied correctly in all relevant dataframes
CONN_LMM_long_old_combined$ROI <- factor(CONN_LMM_long_old_combined$ROI, levels = roi_order)
network_predictions_s_network$ROI <- factor(network_predictions_s_network$ROI, levels = roi_order)
mle_predicted_seg$ROI <- factor(mle_predicted_seg$ROI, levels = roi_order)
p_value_s_overall$ROI <- factor(p_value_s_overall$ROI, levels = roi_order)

# ** Plot for Between-Network Connectivity**
seg_plot <- ggplot(CONN_LMM_long_old_combined, aes(x = Age, y = seg, group = SubNum, color = ROI)) +
  geom_point(size = 2) +  # Slightly larger points for visibility
  geom_line(alpha = 0.6) +  

  # Regression lines for each network (colored)
  geom_line(data = network_predictions_s_network, 
            aes(x = Age, y = Segregation, color = ROI, group = ROI), 
            inherit.aes = FALSE, size = 1.2) +  

  # Overall Mean regression line in **red**
  geom_line(data = mle_predicted_seg, aes(x = Age, y = seg), 
            inherit.aes = FALSE, color = "black", size = 1.2) +

  # Facet by network
  facet_wrap(~ROI, scales = "free_y", ncol = 3, drop = FALSE) +  
  scale_color_manual(values = roi_colors) +  

  # Improve theme
  theme_minimal(base_size = 14) +  
  labs(x = "Age", y = "Network Segregation") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", family = "Times New Roman"),  
    axis.title.x = element_text(size = 14, face = "bold", family = "Times New Roman"),  
    axis.title.y = element_text(size = 14, face = "bold", family = "Times New Roman"),  
    axis.text = element_text(size = 10),  
    strip.text = element_text(size = 12, face = "bold", family = "Times New Roman"),  
    legend.position = "none"
  ) +  

  # P-value annotation for Overall Mean
  geom_text(data = p_value_s_overall, 
          aes(x = max(CONN_LMM_long_old_combined$Age) - 5,  
              y = min(CONN_LMM_long_old_combined$seg[CONN_LMM_long_old_combined$ROI == "Overall Mean"], na.rm = TRUE) + 0.1,  
              label = p_label_s), 
          inherit.aes = FALSE, size = 4, fontface = "italic", color = "black")

ggsave("Fig7_segregation_change.tiff", plot = seg_plot, width = 17.6, height = 13.5, units = "cm", dpi = 300)


```


```{r DONT NEED THAT: cross-sectional age on Within-network connectivity}
library(ggbreak) 

# Fit the linear model for cross-sectional effect

tt <- summary(model_w_cross)$tTable
p_val <- tt["Btwn_cAge", "p-value"]
p_label <- sprintf("p = %.4f", p_val)


# p_val <- summary(model_w_cross)$coefficients$fixed["Btwn_cAge", "Pr(>|t|)"]
# p_label <- sprintf("p = %.4f", p_val)

# Generate predicted line
age_seq <- seq(min(CONN_LMM_Y_W1$Age), max(CONN_LMM_Y_W1$Age), length.out = 1000)
slope <- fixef(model_w_cross)["Btwn_cAge"]
intercept <- fixef(model_w_cross)["(Intercept)"]

predicted_df <- data.frame(
  Age = age_seq,
  within = intercept + slope * (age_seq - 65)  # Btwn_cAge is Age - 65
)

# Plot the data and regression line

p  <- ggplot(CONN_LMM_Y_W1, aes(x = Age, y = within)) +
  geom_point(alpha = 0.4, size = 1, color = "#6CB0D6") +  # Light blue points
  geom_smooth(data = predicted_df, aes(x = Age, y = within), 
              method = "loess", se = FALSE, color = "red", size = 1.1) +  # Smoothed red line
  annotate("text", 
           x = max(CONN_LMM_Y_W1$Age) - 5, 
           y = min(CONN_LMM_Y_W1$within, na.rm = TRUE) + 0.1, 
           label = sprintf("p = %.4f", p_val),
           size = 5, fontface = "italic", color = "red", 
           family = "Times New Roman") +  # P-value text
  labs(
    title = "Cross-Sectional Effect of Age on Within-Network Connectivity",
    x = "Age",
    y = "Within-network Connectivity"
  ) +
  theme_minimal(base_size = 14) +  
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", family = "Times New Roman"),
    axis.title.x = element_text(size = 12, face = "bold", family = "Times New Roman"),
    axis.title.y = element_text(size = 12, face = "bold", family = "Times New Roman"),
    axis.text = element_text(size = 12, family = "Times New Roman")
  )

ggsave("within_network_connectivity_cross_sectional.png", plot = p, width = 9, height = 6, dpi = 300)

```


```{r Model 1 - THIS WORK: within Now Plot the cross-sectional effect by network}
library(dplyr)
library(ggplot2)

# Step 1: Filter the data
CONN_LMM_Y_W1 <- CONN_LMM_long %>%
  filter(Type %in% c("Young", "W1"))

# Step 2: Rename and order ROI levels
CONN_LMM_Y_W1$ROI <- ifelse(CONN_LMM_Y_W1$ROI == "global", "Overall Mean", CONN_LMM_Y_W1$ROI)

CONN_LMM_Y_W1 <- CONN_LMM_Y_W1 %>%
  mutate(ROI_ordered = factor(ROI, levels = c(
    "DMN", "SMN", "SAL", "DAN", "FP", "LAN", "VIS", "CRB", "Overall Mean"
  )))

# Step 3: Global prediction using model_w_cross (assumes this model exists)
age_seq <- seq(min(CONN_LMM_Y_W1$Age), max(CONN_LMM_Y_W1$Age), length.out = 100)
Btwn_cAge_seq <- age_seq - 65

pred_df_global <- data.frame(
  Age = age_seq,
  Btwn_cAge = Btwn_cAge_seq,
  effect_DMN_global = 0,
  effect_SMN_global = 0,
  effect_SAL_global = 0,
  effect_DAN_global = 0,
  effect_FP_global = 0,
  effect_LAN_global = 0,
  effect_CRB_global = 0
)

coefs <- fixef(model_w_cross)
slope <- coefs["Btwn_cAge"]
intercept <- coefs["(Intercept)"]

# Generate predicted within-network connectivity
predicted_df <- data.frame(
  Age = age_seq,
  within = intercept + slope * (age_seq - 65)  # Centered on age 65
)

# Initialize list to store predictions per ROI
pred_lines <- list()

# Loop through each ROI
for (roi in roi_list) {
  # Subset data for the current ROI
  subset_data <- CONN_LMM_Y_W1 %>% filter(ROI == roi)

  # Check for sufficient data
  if (nrow(subset_data) > 2) {
    # Fit cross-sectional model
    model <- lm(within ~ Btwn_cAge, data = subset_data)
    
    # Generate prediction frame
    pred_df <- data.frame(
      Age = age_seq,
      Btwn_cAge = Btwn_cAge_seq
    )
    
    # Predict
    pred_df$fit_value <- predict(model, newdata = pred_df)
    pred_df$ROI <- roi
    
    # Store
    pred_lines[[roi]] <- pred_df
  }
}

# Combine all into one dataframe
pred_lines <- bind_rows(pred_lines)

# Step 4: Combine with pred_lines
pred_df_global <- predicted_df %>%
  select(Age, within) %>%
  rename(fit_value = within) %>%
  mutate(ROI = "Overall Mean")

# Combine all ROI-specific predictions
pred_lines <- bind_rows(pred_lines) %>%
  filter(ROI != "global") %>%    # If needed
  select(Age, fit_value, ROI)    # Drop Btwn_cAge and keep only what's needed


pred_lines_with_global <- rbind(pred_lines, pred_df_global)

# Step 5: Extract p-value from model_w_cross
tt <- summary(model_w_cross)$tTable
p_val <- tt["Btwn_cAge", "p-value"]
p_label <- sprintf("p = %.4f", p_val)


# Step 5.5 adding the order: 

CONN_LMM_Y_W1$ROI_ordered <- factor(CONN_LMM_Y_W1$ROI, levels = roi_order)
pred_lines_with_global$ROI_ordered <- factor(pred_lines_with_global$ROI, levels = roi_order)

# Step 6: Plot with p-value only on "Overall Mean"
p <- ggplot(CONN_LMM_Y_W1, aes(x = Age, y = within, color = ROI)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_line(data = pred_lines_with_global, aes(x = Age, y = fit_value, group = ROI), size = 1.1) +
  facet_wrap(~ROI_ordered, scales = "free_y", ncol = 3) +
  scale_color_manual(values = roi_colors) + 
  # scale_color_manual(values = c(
  # "DMN" = "#66c2a5",
  # "SMN" = "#fc8d62",
  # "SAL" = "#8da0cb",
  # "DAN" = "#e78ac3",
  # "FP"  = "#a6d854",
  # "LAN" = "#ffd92f",
  # "VIS" = "#e5c494",
  # "CRB" = "#b3b3b3",
  # "Overall Mean" = "red")) + 
  theme_minimal(base_size = 13) +
  labs(
    x = "Age", 
    y = "Within-Network Connectivity") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16, family = "Times New Roman"),
    axis.title = element_text(face = "bold", size = 14, family = "Times New Roman"),
    axis.text = element_text(size = 10),
    strip.text = element_text(size = 12, face = "bold", family = "Times New Roman"),
    legend.position = "none"
  ) +
  geom_text(
  data = data.frame(
    ROI_ordered = factor("Overall Mean", levels = roi_order), 
    Age = 70,  # Adjusted x-position
    within = 0.3,  # Adjusted y-position
    label = p_label
  ),
  aes(x = Age, y = within, label = label),
  inherit.aes = FALSE,
  size = 4,
  family = "Times New Roman",
  color = "black",
  fontface = "italic"
)

# Adding FP network significant results 
# Compute FP p-value (cross-sectional: within ~ Btwn_cAge)
fp_dat <- CONN_LMM_Y_W1 %>% filter(ROI == "FP")
m_fp   <- lm(within ~ Btwn_cAge, data = fp_dat)

p_fp   <- summary(m_fp)$coefficients["Btwn_cAge", "Pr(>|t|)"]
lab_fp <- ifelse(p_fp < 1e-4, "p < .0001", sprintf("p = %.4f", p_fp))

# Place the label nicely inside the FP panel
x_pos <- max(fp_dat$Age, na.rm = TRUE) - 0.05 * diff(range(fp_dat$Age, na.rm = TRUE))
y_pos <- min(fp_dat$within, na.rm = TRUE) + 0.10 * diff(range(fp_dat$within, na.rm = TRUE))

annot_fp <- data.frame(
  ROI_ordered = factor("FP", levels = roi_order),
  Age   = Inf,      # right edge of the FP panel
  within = -Inf,    # bottom edge of the FP panel
  label = lab_fp
)

p <- p +
  geom_text(
    data = annot_fp,
    aes(x = Age, y = within, label = label),
    inherit.aes = FALSE,
    hjust = 1.02,    # nudge left from the right edge
    vjust = -0.6,    # nudge up from the bottom edge
    size = 4,
    family = "Times New Roman",
    fontface = "italic",
    color = "red"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.06)))


ggsave("Fig.2_within_network_connectivity_cross_sectional.tiff", plot = p, width = 17.6, height = 13.5, units = "cm", dpi = 300)

```

```{r Modle 2 = plot cross-sectional age effect on between-network connectivity}
library(dplyr)
library(ggplot2)

# Step 1: Filter the data
CONN_LMM_Y_W1 <- CONN_LMM_long %>%
  filter(Type %in% c("Young", "W1"))

# Step 2: Rename and order ROI levels
CONN_LMM_Y_W1$ROI <- ifelse(CONN_LMM_Y_W1$ROI == "global", "Overall Mean", CONN_LMM_Y_W1$ROI)

CONN_LMM_Y_W1 <- CONN_LMM_Y_W1 %>%
  mutate(ROI_ordered = factor(ROI, levels = c(
    "DMN", "SMN", "SAL", "DAN", "FP", "LAN", "VIS", "CRB", "Overall Mean"
  )))

# Step 3: Global prediction using model_w_cross (assumes this model exists)
age_seq <- seq(min(CONN_LMM_Y_W1$Age), max(CONN_LMM_Y_W1$Age), length.out = 100)
Btwn_cAge_seq <- age_seq - 65

pred_df_global <- data.frame(
  Age = age_seq,
  Btwn_cAge = Btwn_cAge_seq,
  effect_VIS_global = 0,
  effect_SMN_global = 0,
  effect_SAL_global = 0,
  effect_DAN_global = 0,
  effect_FP_global = 0,
  effect_LAN_global = 0,
  effect_CRB_global = 0
)

coefs <- fixef(model_b_cross)
slope <- coefs["Btwn_cAge"]
intercept <- coefs["(Intercept)"]

# Generate predicted within-network connectivity
predicted_df <- data.frame(
  Age = age_seq,
  between = intercept + slope * (age_seq - 65)  # Centered on age 65
)

# Initialize list to store predictions per ROI
pred_lines <- list()

# Loop through each ROI
for (roi in roi_list) {
  # Subset data for the current ROI
  subset_data <- CONN_LMM_Y_W1 %>% filter(ROI == roi)

  # Check for sufficient data
  if (nrow(subset_data) > 2) {
    # Fit cross-sectional model
    model <- lm(between ~ Btwn_cAge, data = subset_data)
    
    # Generate prediction frame
    pred_df <- data.frame(
      Age = age_seq,
      Btwn_cAge = Btwn_cAge_seq
    )
    
    # Predict
    pred_df$fit_value <- predict(model, newdata = pred_df)
    pred_df$ROI <- roi
    
    # Store
    pred_lines[[roi]] <- pred_df
  }
}

# Combine all into one dataframe
pred_lines <- bind_rows(pred_lines)

# Step 4: Combine with pred_lines
pred_df_global <- predicted_df %>%
  select(Age, between) %>%
  rename(fit_value = between) %>%
  mutate(ROI = "Overall Mean")

# Combine all ROI-specific predictions
pred_lines <- bind_rows(pred_lines) %>%
  filter(ROI != "global") %>%    # If needed
  select(Age, fit_value, ROI)    # Drop Btwn_cAge and keep only what's needed


pred_lines_with_global <- rbind(pred_lines, pred_df_global)

# Step 5: Extract p-value from model_w_cross
tt <- summary(model_b_cross_int)$tTable
p_val <- tt["Btwn_cAge", "p-value"]
p_label <- sprintf("p = %.4f", p_val)

# Step 5.5 adding the order: 

CONN_LMM_Y_W1$ROI_ordered <- factor(CONN_LMM_Y_W1$ROI, levels = roi_order)
pred_lines_with_global$ROI_ordered <- factor(pred_lines_with_global$ROI, levels = roi_order)

# Step 6: Plot with p-value only on "Overall Mean"
p <- ggplot(CONN_LMM_Y_W1, aes(x = Age, y = between, color = ROI)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_line(data = pred_lines_with_global, aes(x = Age, y = fit_value, group = ROI), size = 1.1) +
  facet_wrap(~ROI_ordered, scales = "free_y", ncol = 3) +
  scale_color_manual(values = roi_colors) + 
  # scale_color_manual(values = c(
  # "DMN" = "#66c2a5",
  # "SMN" = "#fc8d62",
  # "SAL" = "#8da0cb",
  # "DAN" = "#e78ac3",
  # "FP"  = "#a6d854",
  # "LAN" = "#ffd92f",
  # "VIS" = "#e5c494",
  # "CRB" = "#b3b3b3",
  # "Overall Mean" = "red")) + 
  theme_minimal(base_size = 13) +
  labs(
    x = "Age", 
    y = "Between-Network Connectivity") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16, family = "Times New Roman"),
    axis.title = element_text(face = "bold", size = 14, family = "Times New Roman"),
    axis.text = element_text(size = 10),
    strip.text = element_text(size = 12, face = "bold", family = "Times New Roman"),
    legend.position = "none"
  ) +
  geom_text(
  data = data.frame(
    ROI_ordered = factor("Overall Mean", levels = roi_order), 
    Age = 70,  # Adjusted x-position
    between = 0.135,  # Adjusted y-position
    label = p_label
  ),
  aes(x = Age, y = between, label = label),
  inherit.aes = FALSE,
  size = 4,
  family = "Times New Roman",
  fontface = "italic",
  # color = "red"
)

ggsave("Figure3_Between_network_connectivity_cross_sectional.tiff", plot = p, width = 17.6, height = 13.5, units = "cm", dpi = 300)

```

```{r Model 2 = Cross-sectional effect of age on network segregation}
library(dplyr)
library(ggplot2)

# Step 1: Filter the data
CONN_LMM_Y_W1 <- CONN_LMM_long %>%
  filter(Type %in% c("Young", "W1"))

# Step 2: Rename and order ROI levels
CONN_LMM_Y_W1$ROI <- ifelse(CONN_LMM_Y_W1$ROI == "global", "Overall Mean", CONN_LMM_Y_W1$ROI)

CONN_LMM_Y_W1 <- CONN_LMM_Y_W1 %>%
  mutate(ROI_ordered = factor(ROI, levels = c(
    "DMN", "SMN", "SAL", "DAN", "FP", "LAN", "VIS", "CRB", "Overall Mean"
  )))

# Step 3: Global prediction using model_w_cross (assumes this model exists)
age_seq <- seq(min(CONN_LMM_Y_W1$Age), max(CONN_LMM_Y_W1$Age), length.out = 100)
Btwn_cAge_seq <- age_seq - 65

pred_df_global <- data.frame(
  Age = age_seq,
  Btwn_cAge = Btwn_cAge_seq,
  effect_VIS_global = 0,
  effect_SMN_global = 0,
  effect_SAL_global = 0,
  effect_DAN_global = 0,
  effect_FP_global = 0,
  effect_LAN_global = 0,
  effect_CRB_global = 0
)

coefs <- fixef(model_s_cross_int)
slope <- coefs["Btwn_cAge"]
intercept <- coefs["(Intercept)"]

# Generate predicted within-network connectivity
predicted_df <- data.frame(
  Age = age_seq,
  seg = intercept + slope * (age_seq - 65)  # Centered on age 65
)

# Initialize list to store predictions per ROI
pred_lines <- list()

# Loop through each ROI
for (roi in roi_list) {
  # Subset data for the current ROI
  subset_data <- CONN_LMM_Y_W1 %>% filter(ROI == roi)

  # Check for sufficient data
  if (nrow(subset_data) > 2) {
    # Fit cross-sectional model
    model <- lm(seg ~ Btwn_cAge, data = subset_data)
    
    # Generate prediction frame
    pred_df <- data.frame(
      Age = age_seq,
      Btwn_cAge = Btwn_cAge_seq
    )
    
    # Predict
    pred_df$fit_value <- predict(model, newdata = pred_df)
    pred_df$ROI <- roi
    
    # Store
    pred_lines[[roi]] <- pred_df
  }
}

# Combine all into one dataframe
pred_lines <- bind_rows(pred_lines)

# Step 4: Combine with pred_lines
pred_df_global <- predicted_df %>%
  select(Age, seg) %>%
  rename(fit_value = seg) %>%
  mutate(ROI = "Overall Mean")

# Combine all ROI-specific predictions
pred_lines <- bind_rows(pred_lines) %>%
  filter(ROI != "global") %>%    # If needed
  select(Age, fit_value, ROI)    # Drop Btwn_cAge and keep only what's needed


pred_lines_with_global <- rbind(pred_lines, pred_df_global)

# Step 5: Extract p-value from model_w_cross
tt <- summary(model_s_cross_int)$tTable
p_val <- tt["Btwn_cAge", "p-value"]
p_label <- sprintf("p = %.4f", p_val)

# Step 5.5 adding the order: 

CONN_LMM_Y_W1$ROI_ordered <- factor(CONN_LMM_Y_W1$ROI, levels = roi_order)
pred_lines_with_global$ROI_ordered <- factor(pred_lines_with_global$ROI, levels = roi_order)

# Step 6: Plot with p-value only on "Overall Mean"
p <- ggplot(CONN_LMM_Y_W1, aes(x = Age, y = seg, color = ROI)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_line(data = pred_lines_with_global, aes(x = Age, y = fit_value, group = ROI), size = 1.1) +
  facet_wrap(~ROI_ordered, scales = "free_y", ncol = 3) +
  scale_color_manual(values = roi_colors) + 
  # scale_color_manual(values = c(
  # "DMN" = "#66c2a5",
  # "SMN" = "#fc8d62",
  # "SAL" = "#8da0cb",
  # "DAN" = "#e78ac3",
  # "FP"  = "#a6d854",
  # "LAN" = "#ffd92f",
  # "VIS" = "#e5c494",
  # "CRB" = "#b3b3b3",
  # "Overall Mean" = "red")) + 
  theme_minimal(base_size = 13) +
  labs(
    x = "Age", 
    y = "Network Segregation") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16, family = "Times New Roman"),
    axis.title = element_text(face = "bold", size = 14, family = "Times New Roman"),
    axis.text = element_text(size = 10),
    strip.text = element_text(size = 12, face = "bold", family = "Times New Roman"),
    legend.position = "none"
  ) +
  geom_text(
  data = data.frame(
    ROI_ordered = factor("Overall Mean", levels = roi_order), 
    Age = 70,  # Adjusted x-position
    seg = 0.7,  # Adjusted y-position
    label = p_label
  ),
  aes(x = Age, y = seg, label = label),
  inherit.aes = FALSE,
  size = 4,
  family = "Times New Roman",
  fontface = "italic",
  # color = "red"
)

# ---- Add SAL p-value to the SAL panel ----
sal_dat <- CONN_LMM_Y_W1 %>% dplyr::filter(ROI == "SAL")
m_sal   <- lm(seg ~ Btwn_cAge, data = sal_dat)

p_sal   <- summary(m_sal)$coefficients["Btwn_cAge", "Pr(>|t|)"]
lab_sal <- ifelse(p_sal < 1e-4, "p < .0001", sprintf("p = %.4f", p_sal))
col_sal <- ifelse(p_sal < 0.05, "red", "black")  # highlight if significant

annot_sal <- data.frame(
  ROI_ordered = factor("SAL", levels = roi_order),
  Age   = Inf,     # anchor to right edge of SAL panel
  seg   = -Inf,    # anchor to bottom edge
  label = lab_sal
)

p <- p +
  geom_text(
    data = annot_sal,
    aes(x = Age, y = seg, label = label),
    inherit.aes = FALSE,
    hjust = 1.02,      # nudge left from right edge
    vjust = -0.6,      # nudge up from bottom
    size = 4,
    family = "Times New Roman",
    fontface = "italic",
    color = col_sal
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_continuous(expand = expansion(mult = c(0.06, 0.06)))


ggsave("Fig4_Network_segregation_cross_sectional.tiff", plot = p, width = 17.6, height = 13.5, units = "cm", dpi = 300)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
